\documentclass[MSc,12pt]{wsuthesis}

%\usepackage[hdvipdfm]{graphics}
\usepackage[table]{xcolor}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{color}
\usepackage{pstricks}
\usepackage{setspace}
\usepackage{url}
\usepackage{listings}
\usepackage{appendix}
\usepackage{lscape}
\usepackage{graphicx}
\usepackage{times}
\usepackage{cite} 
\usepackage{subcaption}
\usepackage{listings} 
\usepackage{url}
\usepackage{changepage}
\usepackage{multirow}

% import figures
%\input{figures/figures.tex}

\pdfpxdimen=1in
\divide\pdfpxdimen by 600
\setlength{\fboxsep}{0pt}
\setlength{\fboxrule}{2px}
\definecolor{wsu-gray}{RGB}{113,113,113}
\definecolor{wsu-light-gray}{RGB}{215,218,219}


\lstnewenvironment{algorithm}[1][] %defines the algorithm listing environment
{   
    \lstset{ %this is the stype
        frame=tB,
        numbers=left, 
        showlines=true,
        numberstyle=\tiny,
        basicstyle=\scriptsize, 
        keywordstyle=\color{black}\bfseries,
        keywords={,input, output, return, datatype, function, in, if, else, 
                   foreach, while, begin, end, for, all, do, out, all, then, is,
                   not, each, }
        numbers=left,
        xleftmargin=.04\textwidth,
        #1 
    }
}{}

\lstnewenvironment{cnc}[1][] %defines the algorithm listing environment
{   
    \lstset{ %this is the stype
        frame=tB,
        numbers=left, 
        showlines=true,
        numberstyle=\tiny,
        basicstyle=\scriptsize, 
        keywordstyle=\color{black}\bfseries,
        keywords={, \$context, \$range, \$initialize, \$finalize, (, ),
        [, ], )), } numbers=left,
        xleftmargin=.04\textwidth,
        #1 
    }
}{}

%\newcommand\fnurl[2]{%
%  \href{#2}{#1}\footnote{\url{#2}}%
%}

\begin{document}

\title{Communication Avoiding Ray Tracing \\ for Exascale Computing}

\author{Ellen Alice Porter}

\submitdate{July 2016}

\centerhdrstrue
\strhdrstrue

\dept{School of Electrical Engineering and Computer Science}

\chair{Robert R. Lewis}

\acknowledgment{
  The completion of this thesis would not have been possible without the help of
  many people.  I would like to express my sincerest gratitude to my advisor,
  Dr. Robert R. Lewis for his guidance and patience over the years.  I would
  also like to acknowledge my thesis committee including Dr. Wayne O.
  Cochran and Dr. Carl H. Hauser for their commitment towards my education. The
  CnC community was essential to the completion of this research, especially
  Kath Knobe and Nick Vrvilo, thank you both.  I must also recognize my
  employer, Battelle, my managers and my colleagues for their support and
  encouragement.  Last but not least, I would like to thank my family,
  especially Mathieu, for their unconditional faith, inspiration and love.
}

\abstract{
  Exascale computers, defined as being capable of performing at least
  one exaFLOP ($10^{18}$ floating point operations per second) are
  anticipated to emerge in the next several years. Reaching this scale
  of computation will require hardware and software changes for
  high-performance computing (HPC). Applications will need to adapt as
  the architecture of supercomputers change. Some studies suggest
  \emph{communication avoiding} algorithms might be the most performant
  design for future systems. This has created an interest in the
  development of scalable visualization algorithms and techniques.

  This thesis explores the impact exascale hardware will have on
  programming models and application design. We look at one
  specific programming model, Concurrent Collections (CnC), and explore how we
  can use it coupled with Intel's ray tracing engine, Embree, to build a
  scalable ray tracing system with an emphasis on communication avoidance and
  extension for exascale. %
}

\dedication{
  \emph{For my mom and dad.}
}

\preface

\chapter{Introduction}
\input{ch1_introduction} % subsections: introduction, exascale, ray tracing
  
\chapter{Previous Work} % or literature review/background?
\input{ch2_previous_work} % subsections: distributed ray tracing, task-based
                          % programming models, cnc model, petri net

\chapter{Design}
\input{ch3_design} 

\chapter{Implementation}
\input{ch4_implementation}
  
\chapter{Evaluation}
\input{ch5_evaluation}

\chapter{Conclusion and Future Work}
\label{future-work-and-conclusions}
\input{ch6_conclusions}
\input{ch7_future_work}


\newpage
\singlespacing
\bibliographystyle{plain}
\bibliography{bibliography}

\appendix
% If you really *must* include source code.
\chapter{Source Code}
\input{appendix.tex}

\end{document}
